\Large\bf New features in FlexibleSUSY since version 2.0

### 2019-03-04 doc/flexiblesusy-new_features.tex

\usepackage[pdftitle={FlexibleSUSY 2.0: Extensions to investigate the phenomenology of
  SUSY and non-SUSY models},
  pdfauthor={Peter Athron, Markus Bach, Dylan Harries, Thomas Kwasnitza, Jae-hyeon Park,
    Dominik Stockinger, Alexander Voigt, Jobst Ziebell},
  pdfkeywords={FlexibleSUSY,supersymmetry,spectrum,generator,MSSM,NMSSM,E6SSM},
  bookmarks=true, linktocpage, colorlinks=true, allbordercolors=white,
  allcolors=fscolor]{hyperref}
\biboptions{sort&compress}
\allowdisplaybreaks

@@

@@

\journal{Computer Physics Communications}
\begin{document}
\begin{frontmatter}
 \vspace*{0.5cm}
 \title{\Large\bf FlexibleSUSY 2.0: Extensions to investigate the phenomenology of SUSY and non-SUSY models}

\author[Monash]{Peter Athron}
\author[dresden]{Markus Bach}
\author[adelaide,prague]{Dylan Harries}
\author[dresden]{Thomas Kwasnitza}
\author[kiasquc]{Jae-hyeon Park}
\author[dresden]{Dominik St\"ockinger}
\author[rwth]{Alexander Voigt\corref{cor1}}
\ead{Alexander.Voigt@physik.rwth-aachen.de}
\cortext[cor1]{Corresponding author}
\author[dresden]{Jobst Ziebell}
\address[Monash]{ARC Centre of Excellence for Particle Physics at
  the Terascale, School of Physics, Monash University, Melbourne,
  Victoria 3800, Australia}
\address[dresden]{Institut f\"ur Kern- und Teilchenphysik,
TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}
\address[adelaide]{ARC Centre of Excellence for Particle Physics at
the Terascale, Department of Physics, The University of Adelaide,
Adelaide, South Australia 5005, Australia}
\address[prague]{Institute of Particle and Nuclear Physics, Faculty of
  Mathematics and Physics, Charles University in Prague, V
  Hole\v{s}ovi\v{c}k\'{a}ch 2, 180 00 Praha 8, Czech Republic}
\address[kiasquc]{Quantum Universe Center,
Korea Institute for Advanced Study,
85 Hoegiro Dongdaemungu,
Seoul 02455, Republic of Korea}
\address[rwth]{Institute for Theoretical Particle Physics and Cosmology, RWTH Aachen University, 52074 Aachen, Germany}

  \begin{abstract}
    We document major new features and improvements of \fs, a
    \mathematica and C++ package with a dependency on the external
    package \sarah, that generates fast and precise spectrum
    generators. The extensions presented here significantly increase
    the generality and capabilities of the \fs package, which already
    works with a wide class of models, while
    maintaining an elegant structure and easy to use interfaces. The
    \fbsm extension makes it possible to also create spectrum
    generators for non-supersymmetric extensions of the Standard
    Model. The \fcpv extension adds the option of complex parameters
    to the spectrum generators, allowing the study of many interesting
    models with new sources of \CP violation. \fmw computes the decay
    of the muon for the generated model and thereby allows \fs to
    predict the mass of the $W$ boson from the input parameters by
    using the more precise electroweak input of $\{ G_F, M_Z, \aem \}$
    instead of $\{ M_W, M_Z, \aem \}$. The \famu extension provides a
    calculator of the anomalous magnetic moment of the muon in any
    model \fs can generate a spectrum for.  \fsas introduces a new
    solver for the boundary value problem which makes use of
    semi-analytic expressions for dimensionful parameters to find
    solutions in models where the classic two-scale solver will not
    work such as the constrained E$_6$SSM\@. \feft is a hybrid
    calculation of the Higgs mass which combines the virtues of both
    effective field theory calculations and fixed-order calculations.
    All of these extensions are included in \fstwo, which is released
    simultaneously with this manual.%% \sarah, an external dependency
    %% providing model dependent information, can be automaically
    %% installed using an installation script distributed with \fs.
  \end{abstract}

@@

\clearpage
\newgeometry{top=2.5cm,left=3cm,right=3cm,bottom=4cm,footskip=3em}
\section*{New version program summary}
\noindent
{\em Program title:} \fs\\[0.5em]
{\em Licensing provisions:} GPLv3\\[0.5em]
{\em Program obtainable from:} \url{http://flexiblesusy.hepforge.org/}\\[0.5em]
{\em Programming language:} C++, Wolfram/Mathematica, FORTRAN, Bourne shell\\[0.5em]
{\em Operating system:} Tested on FreeBSD, Linux, Mac OS X, Windows/Cygwin\\[0.5em]
{\em External routines:} \sarah, Boost library, Eigen library, GNU Scientific Library\\[0.5em]
{\em Optional external routines:} BLAS, Himalaya, LAPACK, LoopTools, SQLite, TSIL\\[0.5em]
{\em Journal reference of previous version:} Comput.Phys.Commun.\ 190 (2015) 139-172\\[0.5em]
{\em Does the new version supersede the previous version?:} yes \\[0.5em]
{\em Reasons for the new version:} Program extension including new models, observables and algorithms\\[0.5em]
{\em Summary of revisions:} Extension to non-supersymmetric models
(\fbsm), complex parameters (\fcpv), prediction of $W$ boson mass from
muon decay (\fmw), calculation of anomalous magnetic moment of the
muon (\famu), semi-analytic boundary value problem solver (\fsas),
improved hybrid Higgs mass calculation (\feft).\\[0.5em]
{\em Nature of problem:}
%
Determining the mass spectrum, mixings and further observables for an arbitrary extension
of the Standard Model, input by the user. The generated code must find
simultaneous solutions to constraints that are specified at two or
more different renormalization scales, which are connected by
renormalization group equations forming a large set of coupled
first-order differential
equations. \\[0.5em]
%
{\em Solution method:}
%
Nested iterative algorithm and numerical
minimization of the Higgs potential.\\[0.5em]
%
{\em Restrictions:}
%
The couplings must remain perturbative at all scales between the
highest and lowest boundary condition.  Tensor-like Lagrangian
parameters of rank 3 are currently not supported.  The automatic
determination of the Standard Model-like gauge and Yukawa couplings is
only supported for models that have the Standard Model gauge group
$SU(3)_C\times SU(2)_L\times U(1)_Y$ as a gauge symmetry group factor.
However, due to the modular nature of the generated code, adapting and
extending it to overcome restrictions in scope is quite straightforward.

@@

\newpage
\section{Introduction}

Popular and well studied new physics extensions of the Standard Model
(SM) are coming under increasing pressure from the searches at the
Large Hadron Collider (LHC) as well as other current experiments.  At
the same time, there remain many outstanding physics problems that are
not solved by the SM and require new physics.  For example, the origin
of dark matter is still unexplained and it requires some new particle
to fit the relic density.  In addition, the gauge structure and
fractional charges of the SM have no explanation, but some
observations hint at the possibility of a grand unified theory (GUT)
from which these can be derived after the breakdown of the GUT gauge
group.  Finally, the stability of the weak scale, which is $17$ orders
of magnitude smaller than the Planck scale, has no explanation within
the SM \cite{Weinberg:1975gm, Weinberg:1979bn,
  Gildener:1976ai,Susskind:1978ms, tHooft:1980xss} and has been the
driving motivation for the construction and study of concepts that go
beyond the Standard Model (BSM), most notably supersymmetry (SUSY) \cite{Miyazawa:1966mfa,Miyazawa:1968zz,Ramond:1971gb,Golfand:1971iw,Gervais:1971ji,Neveu:1971rx,Neveu:1971iv,Volkov:1973ix,Wess:1973kz,Wess:1974tw,Salam:1974yz} which ensures that the quadratic corrections from fermions and bosons in the loop diagrams cancel at all orders in perturbation theory \cite{Dimopoulos:1981au,Witten:1981nf,Dine:1981za,Dimopoulos:1981zb,Kaul:1981hi}.

These issues strongly motivate the development of new ideas and new
models.  However, in principle there are a huge number of models that
can solve some, or all, of these problems and many more that may be
motivated by other principles that have not yet gained widespread
interest in high energy physics (HEP)\@. Previously, most work has
been done on the simplest variants, on scenarios that are easiest to
test and on those that individual researchers consider to be the very
best motivated models. Examples include the Minimal Supersymmetric
Standard Model\footnote{For a review see Ref.\ \cite{Chung:2003fi}.}
(MSSM), scalar singlet dark matter models (SSDM)
\cite{Silveira:1985rk,McDonald:1993ex,Burgess:2000yq}, type II two
Higgs doublet models (THDM-II) \cite{Lee:1973iz,Glashow:1976nt,
  Donoghue:1978cj}, and universal extra dimensions
\cite{Appelquist:2000nn}.  Now, it is becoming increasingly well
motivated to also examine more complicated model variants, explore
scenarios that are calculationally difficult or hard to observe, or
motivated from an entirely new perspective.  Such advanced models can
avoid phenomenological or conceptual difficulties of simpler models,
and might provide explanations for the lack of experimental evidence
for new physics at the LHC.

Faced with this challenge, it is important to reduce
the calculational hurdle required to explore a new model and
look at its phenomenology as much as possible. This makes it easier for
new ideas to get developed and for many more models to
be studied together for much more general conclusions.

\fs \cite{Athron:2014yba,Athron:2016fuq} already made a significant
push in this direction, allowing the automatic creation of a spectrum
generator for a very wide range of supersymmetric models.  \fs uses
\sarah \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta} to obtain \mathematica expressions for model dependent
components, the 2-loop renormalization group equations (RGEs),
1-loop self energies, 1-loop tadpoles, mass matrices and
electroweak symmetry breaking (EWSB) conditions. \fs then translates these
expressions into C++ routines and embeds them inside a code structure
for solving the boundary value problem (BVP)\@.  It also uses some numerical
routines from \softsusy \cite{Allanach:2001kg,Allanach:2013kza} and is
heavily unit-tested against the MSSM and Next-to-MSSM versions of
\softsusy every night to ensure bugs are avoided in updates.

Spectrum generators determine the pole masses and couplings of a
particular model from assumptions about the parameters at some
high scale, such as the grand unification scale, or directly from
on-shell parameters or running parameters\footnote{Usually defined in
  the \MSbar or \DRbar scheme.} given at the new physics scale.
This requires computing self energies, tadpole corrections to the EWSB
conditions and threshold corrections to SM-like gauge and Yukawa
couplings.  In the case of high scale assumptions a BVP must be
solved, which additionally requires integrating the RGEs for the model.
These codes are essential for testing hypotheses about the parameters of
the model, and determining if they lead to phenomenologically viable masses.

Spectrum generators are widely used in studies of supersymmetry, and for the
MSSM and the Next-to-MSSM\footnote{For a review of the model see
  Refs.\ \cite{Ellwanger:2009dp, Maniatis:2009re}.} (NMSSM) there are
a number of them that solve the BVP for particular choices of breaking
mechanism inspired boundary conditions (\softsusy
\cite{Allanach:2001kg,Allanach:2009bv, Allanach:2011de,
  Allanach:2013kza, Allanach:2014nba,
  Allanach:2016rxd,Allanach:2017hcf}, \suspect \cite{Djouadi:2002ze},
\spheno \cite{Porod:2003um,Porod:2011nf}, \isasusy \cite{Baer:1993ae},
\nmspec \cite{Ellwanger:2004xm,Ellwanger:2005dv, Ellwanger:2006rn} and
\suseflav \cite{Chowdhury:2011zr}) and several more that start from \DRbar
parameters at the SUSY scale or on-shell inputs (\FeynHiggs
\cite{Heinemeyer:1998yj, Heinemeyer:1998np, Degrassi:2002fi,
  Frank:2006yh, Hahn:2013ria, Drechsel:2016jdg, Bahl:2016brp,
  Bahl:2017aev}, \NMSSMCalc \cite{Baglio:2013iia, King:2015oxa} and
\CPsuperH \cite{Lee:2003nta, Lee:2007gn, Lee:2012wa}). This means that
the parameter space of the MSSM (and to a lesser extent the NMSSM) has
been extensively explored in a fast and reliable way.

However, for other models few, if any, public software packages exist.
Until recently this meant that one would have to spend a very long
time writing and testing code in order to explore the parameter space
of the model and do phenomenological investigations, and even then
bugs are more likely if one uses a private tool rather than a well
tested public one.  With the recent development of \fs and generated
modules for \spheno from \sarah \cite{Porod:2003um,
  Porod:2011nf,Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta,Athron:2014yba} it is now possible to obtain spectrum
generators in a much wider range of models.  \fs can generate spectrum
generators of various kinds, with boundary conditions at some high
scale or purely with low-energy input.  This push towards calculators
that are not model specific has also happened for other major types of
analyses.  For example, one may create new models in \micromegas
\cite{Belanger:2001fz,
  Belanger:2004yn,Belanger:2006is,Belanger:2008sj,Belanger:2010gh,
  Belanger:2013oya,Belanger:2014vza} using \calchep \cite{Boos:1994xb,
  Belyaev:2012qa} and \lanhep \cite{Semenov:1996es,Semenov:1998eb,
  Semenov:2002jw,Semenov:2008jy, Semenov:2010qt} to study the relic
density and direct and indirect detection of dark matter. \FeynRules
\cite{Christensen:2008py,Alloul:2013bka} can
also be used to generate the Feynman rules after inputting the
Lagrangian, in a similar manner to \sarah and \lanhep. The output of
these codes can be used in matrix element generators, such as \calchep,
\madgraph
\cite{Stelzer:1994ta,Maltoni:2002qb,Alwall:2007st,Alwall:2011uj,
  Alwall:2014hca}, \whizard \cite{Moretti:2001zz,Kilian:2007gr},
\sherpa \cite{Gleisberg:2003xi,Gleisberg:2008ta,Hoeche:2012yf}, \FAFC
\cite{Hahn:1998yk,Hahn:2000kx,Hahn:2006zy,Hahn:2010zi,Hahn:2016ebn}
and \helac \cite{Kanaki:2000ey,Cafarella:2007pc}, with showering and
hadronization handled by the event generators \herwig
\cite{Bahr:2008pv, Bellm:2015jjp}, \pythia
\cite{Bengtsson:1987kr,Sjostrand:2014zea, Sjostrand:2006za} and
\sherpa. Results from collider experiments can be applied to different
BSM models using re-interpretation tools, with some examples being \higgsbounds
\cite{Bechtle:2008jh,Bechtle:2011sb,Bechtle:2013wla, Bechtle:2015pma},
\higgssignals \cite{HiggsSignals}, \checkmate
\cite{Drees:2013wra,Dercks:2016npn}, \smodels \cite{Kraml:2013mwa,
  Kraml:2014sna,Ambrogi:2017neo}, \fastlim \cite{Papucci:2014rja},
\madanalysis\ \cite{Conte:2012fm,Conte:2014zja,Dumont:2014tja}, and
the native re-interpretation of \colliderbit \cite{Balazs:2017moi}.
Finally, very recently \GAMBIT \cite{Athron:2017ard} has been released
which uses auto-generated code from \fs \cite{Workgroup:2017bkh} and
\micromegas \cite{Workgroup:2017lvb}, along with many other packages,
to perform global fits of a user implemented BSM model.

Indeed \fs has been used extensively to study new physics models, on
its own or in combination with some of the codes mentioned above, in a
large number of physics studies, see, e.g.,
Refs.\ \cite{Diessner:2014ksa,
  Athron:2014pua,Athron:2015tsa,Diessner:2015yna,Staub:2015aea,Athron:2015vxg,
  Diessner:2015iln,Bagnaschi:2015pwa,Staub:2016dxq,Athron:2016fuq,
  deFlorian:2016spz,Athron:2016gor,Athron:2016qqb,Drechsel:2016htw,
  Bagnaschi:2017xid,Athron:2017yua,Athron:2017kgt,Athron:2017qdc,
  Workgroup:2017bkh,Chakravarty:2017hcy,Harlander:2017kuc,Ellis:2017hdw}.
Nonetheless, in the first release of \fs there were still a number of
limitations that restricted the models and phenomenology that could be studied
and the precision of the calculations.

In this paper, we document extensions to \fs, all now available in
version 2.0, that make a substantial push further in expanding the
number of models that can be explored and the observables that can be
calculated within them.
Now the list of models that can be investigated with
\fstwo also includes non-supersymmetric models (\fbsm), those with
complex parameters (\fcpv) and, with the new BVP solver
\fsas, constrained versions of certain non-minimal SUSY models like the NMSSM
\cite{Ellwanger:2009dp, Maniatis:2009re,Ellwanger:1993xa,Elliott:1994ht,
  King:1995vk,Ellwanger:1995ru,Ellwanger:1996gw} and the Exceptional
Supersymmetric Standard Model (\ESSM)
\cite{King:2005jy,King:2005my,Athron:2010zz, Athron:2009ue,Athron:2009bs} that
cannot be solved using the two-scale BVP solver
\cite{Barger:1993gh} approach implemented in public spectrum
generators.
%
In addition, \fstwo extends its repertoire of calculations.
The user may now calculate the mass of the $W$ boson as a
prediction of the model (\fmw) with the Fermi
constant treated as an input, as well as the anomalous magnetic
moment of the muon (\famu) and some electric dipole moments as
an application of \fcpv. Furthermore, we document and update our \feft
calculation of the Higgs pole mass, the physics of which has been discussed
in Ref.~\cite{Athron:2016fuq}.  Besides these new physics features, we
also document the new \mathematica interface.

In \secref{sec:download} we give a quick start guide explaining how to
download and compile the code using basic options.  In
\secref{sec:Mathematica_interface} we present the \mathematica
interface of \fs's spectrum generators and in
\secref{sec:model_file_extensions} we describe the new model file
extensions.  The
following Sections \ref{sec:fbsm}--\ref{sec:feft} document
each of the major extensions to \fs in a
modular fashion so that readers
can easily skip to a particular section.  These extensions are
as follows.  \fbsm, documented in \secref{sec:fbsm}, allows \fs to work in
non-SUSY models as well as SUSY models; the calculation of the
anomalous magnetic moment of the muon is
given in \secref{sec:famu}; \fcpv, described in
\secref{sec:fcpv}, introduces complex parameters to \fs so that \CP
violating effects may be considered; \fmw calculates the
muon decay, allowing \fs to predict $M_W$, and
is documented in \secref{sec:fmw}; a new BVP solver \fsas is
introduced in \secref{sec:fsas} and finally the hybrid Higgs mass
calculation, \feft, which combines the benefits of both effective field theory
and fixed-order calculations is documented in \secref{sec:feft}.
After this, we briefly summarize remaining limitations of the code in
\secref{sec:limitations} before adding concluding remarks in
\secref{sec:conclusions}.
%
In the appendices we provide a reference of all input parameters and
configuration options of \fstwo as well as new models.

\section{Quick start}
\label{sec:download}

\subsection{Requirements}

The build process of a custom spectrum generator using \fs requires
the following:
%
\begin{itemize}
\item \mathematica, version 7 or higher
\item \sarah, version 4.11.0 or higher [\url{http://sarah.hepforge.org}]
\item C++11 compatible compiler (\texttt{g++} 4.8.5 or higher,
  \texttt{clang++} 3.8 or higher, \texttt{icpc} 15.0 or higher)
\item FORTRAN compiler (\texttt{gfortran}, \texttt{ifort} etc.)
\item Eigen library, version 3.1 or higher
  [\url{http://eigen.tuxfamily.org}]
\item Boost library, version 1.37.0 or higher
  [\url{http://www.boost.org}]
\item GNU scientific library [\url{http://www.gnu.org/software/gsl}]
\end{itemize}
%
Optional:
%
\begin{itemize}
\item an implementation of LAPACK [\url{http://www.netlib.org/lapack}]
  such as ATLAS [\url{http://math-atlas.sourceforge.net}] or
  Intel Math Kernel Library\\\ [\url{http://software.intel.com/intel-mkl}]
\item LoopTools, version 2.8 or higher
  [\url{http://www.feynarts.de/looptools}]
\end{itemize}

\subsection{Downloading \fs and generating a first spectrum generator}
\label{sec:quick-start-cmssm}
\fs is available as gzipped tarball on
[\url{http://flexiblesusy.hepforge.org}] or
[\url{https://github.com/FlexibleSUSY/FlexibleSUSY}] under the version
control system \texttt{git}.
%
To download and install \fstwo as a gzipped tarball run at the command
line:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.1.tar.gz
$ tar -xf FlexibleSUSY-2.0.1.tar.gz
$ cd FlexibleSUSY-2.0.1
\end{lstlisting}%% $
%
\fstwo is distributed with a huge selection of predefined ``models'', 
including several MSSM and NMSSM scenarios such as the CMSSM (called
\modelname{CMSSM}), high-scale SUSY and split-SUSY (called \HSSUSY and
\modelname{SplitMSSM}; these  models have been created for
Ref.~\cite{Bagnaschi:2015pwa}), the semi-constrained and
fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}). The
distribution also contains BSM models such as the
$R$-symmetric MSSM (\modelname{MRSSM}; for a definition of the model
see Ref.~\cite{Kribs:2007ac}), the NUHM \ESSM
(\modelname{E6SSM}, see \cite{King:2005jy,King:2005my,Athron:2010zz,
  Athron:2009ue,Athron:2009bs,Athron:2007en}) and the two-Higgs 
doublet model type II
(\modelname{THDMII}).  See the 
contents of \code{model_files/} for all predefined model files.

A spectrum generator for any of these models can be built with just
three commands.  For example, the CMSSM spectrum generator can be
created by running the following shell commands:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=CMSSM
$ ./configure --with-models=CMSSM
$ make
\end{lstlisting}%% $
%
The \code{createmodel} command creates the model directory
\code{models/CMSSM/} where the code will be generated and adds the
CMSSM model file and an example SLHA input file to it.  The \code{configure}
script sets up the \code{Makefile}, checking the system requirements
and dependencies.  For more options see \code{./configure
  --help}.  The last command creates the code for the spectrum
generator of this model and compiles it.
%
This generated spectrum generator can then be run from the command
line as:
%
\begin{lstlisting}[language=bash]
$ cd models/CMSSM
$ ./run_CMSSM.x --slha-input-file=LesHouches.in.CMSSM \
                --slha-output-file=LesHouches.out.CMSSM
\end{lstlisting}
%
The spectrum generator reads the CMSSM input parameters from the SLHA input file
\code{LesHouches.in.CMSSM} and first solves a BVP to find a
set of running parameters at the SUSY scale that are consistent with all
boundary conditions specified in the model file, and then calculates the pole
masses, mixing matrices and potentially further observables.
The mass spectrum etc.\ obtained in this way is output
in SLHA format \cite{Skands:2003cj,Allanach:2008qq} as
\code{LesHouches.out.CMSSM}.  See \code{./run_CMSSM.x --help} for more
options.

\fs also provides a \mathematica interface, introduced in version 1.7.0,
to call the generated spectrum generators.  For each spectrum
generator, an example \mathematica script named
\code{models/<model>/run_<model>.m} is created for illustration.  For
example, the \modelname{CMSSM} spectrum generator can be called from within
\mathematica like this:
%
\begin{lstlisting}
Get["models/CMSSM/CMSSM_librarylink.m"];

handle = FSCMSSMOpenHandle[
    fsModelParameters -> {
        m0 -> 125,
        m12 -> 500,
        TanBeta -> 10,
        SignMu -> 1,
        Azero -> 0
    }
];

spectrum = FSCMSSMCalculateSpectrum[handle];
FSCMSSMCloseHandle[handle];

Print[spectrum];
\end{lstlisting}
%
Execute \code{?FSCMSSMOpenHandle} for a list of all allowed options and
input parameters.  In the example above, the \code{spectrum} variable
contains the mass spectrum and the running parameters in the form of a
list of replacement rules:
%
\begin{lstlisting}
{CMSSM -> {
   Pole[M[Glu]] -> 1147.35,
   Pole[M[Sd]] -> {957.993, 997.56, 1000.49, 1000.5, 1045.93, 1045.94},
   Pole[M[Sv]] -> {350.753, 351.913, 351.917},
   Pole[M[Su]] -> {796.653, 1002.67, 1003.96, 1005.06, 1043.07, 1043.07},
   Pole[M[Se]] -> {222.916, 229.983, 230.008, 360.842, 360.846, 361.978},
   Pole[M[hh]] -> {114.836, 713.119},
   Pole[M[Ah]] -> {88.5864, 712.848},
   Pole[M[Hpm]] -> {77.2642, 717.628},
   Pole[M[Chi]] -> {204.054, 385.012, 629.649, 643.612},
   Pole[M[Cha]] -> {385.017, 643.924},
   Pole[M[VWm]] -> 80.3935, ...}
}
\end{lstlisting}
%
More details about the \mathematica interface as well as a neat
example of running \HSSUSY through it can be found in Sections
\ref{sec:Mathematica_interface} and \ref{sec:application_HSSUSY},
respectively.

\subsection{Spectrum generators for alternative models}
\label{sec:quick-start-alternative-models}
If the user instead wants to create a spectrum generator for a model
for which there is no pre-existing model file distributed in
\fs, then the model file can be written. Before a \fs
model file can be written for the spectrum generator, there must exist
\sarah model files, which \fs uses to obtain model dependent
information. \sarah also comes with many pre-defined models, but if an
appropriate model is not available, the users may create their own
\sarah model files and add them to the directory
\code{sarah/<modelname>/} in \fs.  For the writing of a \sarah model
file we refer the reader to the extensive \sarah documentation, for
example Refs.\
\cite{Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta,Athron:2014yba}.

Creating a new \fs model file is straightforward. Full details are
given in the original \fs manual Ref.~\cite{Athron:2014yba}.
Here we just repeat a basic example
in the context of the NMSSM that illustrates the main points:
%
The semi-constrained NMSSM (\modelname{NMSSM}) distributed in \fs has
all the soft-breaking trilinear scalar couplings set to a unified
$\azero$ at the GUT scale.  However, often $A_\lambda$ and $A_\kappa$ are
taken to be non-universal in semi-constrained variants of the NMSSM
since the non-universality of the soft singlet mass already violates
the standard universality assumptions of constrained
models.\footnote{See \secref{sec:fsas} for a new approach that
  allows the fully constrained NMSSM to be solved.} To allow separate
values for $A_\lambda$ and $A_\kappa$ at the GUT scale the \fs model
file \code{model_files/NMSSM/FlexibleSUSY.m.in} should be changed from
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

HighScaleInput = {
  ...
  {T[\[Kappa]], Azero \[Kappa]},
  {T[\[Lambda]], Azero LambdaInput}
   ...
};
\end{lstlisting}
%
into
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput},
           {63, ALambdaInput},
           {64, AKappaInput} };

HighScaleInput = {
  ...
  {T[\[Kappa]], AKappaInput \[Kappa]},
  {T[\[Lambda]], ALambdaInput LambdaInput},
  ...
};
\end{lstlisting}
%
The GUT scale values of $A_\lambda$ and $A_\kappa$ can then be specified in the SLHA input file in the \code{EXTPAR} block by entries $63$ and $64$,
%
\begin{lstlisting}
Block EXTPAR
   61   0.1                  # LambdaInput
   63   -100                 # ALambdaInput
   64   -300                 # AKappaInput

\end{lstlisting}

\section{\mathematica interface}
\label{sec:Mathematica_interface}

The spectrum generators created with \fs can be called from within
\mathematica.  To do that, first the spectrum generator must be built,
as described in \secref{sec:quick-start-cmssm}.  Afterwards, the
provided \mathematica interface functions for the model must be
loaded.  For a given model \code{<model>} this is done by including
the following file in the \mathematica session:
%
\begin{lstlisting}
Get["models/<model>/<model>_librarylink.m"];
\end{lstlisting}
%
This script loads the library
\code{models/<model>/<model>_librarylink.so} into the \mathematica
session (assuming the user is in the \code{FlexibleSUSY/} directory).
Afterwards, the \mathematica interface functions listed in
\tabref{tab:mma_ifce_functions} are available.
%
\begin{table}[tbh]
  \centering
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    Function & Description \\

    \midrule \code{FS<model>OpenHandle[...]} & Takes all model input
    parameters as argument and returns a ``handle'' (a reference) to
    the given parameter point, the associated mass spectrum and observables.\\

    \code{FS<model>CloseHandle[handle]} & Releases the resources
    associated to a given handle.\\

    \code{FS<model>CalculateSpectrum[handle]} & Calculates the mass
    spectrum for a given handle.\\

    \code{FS<model>CalculateObservables[handle]} & Calculates the
    observables for a given handle.\\

    \code{FS<model>ToSLHA[handle]} & Returns a string containing the
    mass spectrum and observables associated to a given handle in SLHA format.\\

    \code{FS<model>Set[handle, ...]} & Changes the input parameters
    associated to a given handle.\\

    \code{FS<model>GetSettings[handle]} & Returns the spectrum generator
    settings (precision goal, loop orders, etc.) associated to
    a given handle.\\

    \code{FS<model>GetSMInputParameters[handle]} & Returns the SM
    input parameters associated to a given handle.\\

    \code{FS<model>GetInputParameters[handle]} & Returns the
    model-specific input parameters associated to a given handle.\\

    \code{FS<model>GetProblems[handle]} & Returns a list of problems
    that occurred when calculating the spectrum.\\

    \code{FS<model>GetWarnings[handle]} & Returns a list of warnings
    that occurred when calculating the spectrum.\\
    \bottomrule
  \end{tabularx}
  \caption{\mathematica interface functions provided for a \fs model with the name \code{<model>}.}
  \label{tab:mma_ifce_functions}
\end{table}
%

>>>>

\usepackage[pdftitle={New features in FlexibleSUSY since version 2.0},
pdfauthor={Peter Athron, Markus Bach, Dylan Harries, Wojciech
  Kotlarski, Thomas Kwasnitza, Jae-hyeon Park, Dominik Stockinger,
  Alexander Voigt, Jobst Ziebell},
pdfkeywords={FlexibleSUSY,supersymmetry,spectrum,generator,MSSM,NMSSM,E6SSM},
bookmarks=true, linktocpage, colorlinks=true, allbordercolors=white,
allcolors=fscolor]{hyperref} \biboptions{sort&compress}
\allowdisplaybreaks

@@

\newcommand{\TSIL}{\texttt{TSIL}\@\xspace}
\newcommand{\LoopTools}{\texttt{LoopTools}\@\xspace}

@@

\journal{Computer Physics Communications}
\begin{document}
\begin{frontmatter}
 \vspace*{0.5cm}
 \title{\Large\bf New features in FlexibleSUSY since version 2.0}

\author[Monash]{Peter Athron}
\author[dresden]{Markus Bach}
\author[adelaide,prague]{Dylan Harries}
\author[dresden]{Wojciech Kotlarski}
\author[dresden]{Thomas Kwasnitza}
\author[kiasquc]{Jae-hyeon Park}
\author[dresden]{Dominik St\"ockinger}
\author[rwth]{Alexander Voigt\corref{cor1}}
\ead{Alexander.Voigt@physik.rwth-aachen.de}
\cortext[cor1]{Corresponding author}
\author[dresden]{Jobst Ziebell}
\address[Monash]{ARC Centre of Excellence for Particle Physics at
  the Terascale, School of Physics, Monash University, Melbourne,
  Victoria 3800, Australia}
\address[dresden]{Institut f\"ur Kern- und Teilchenphysik,
TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}
\address[adelaide]{ARC Centre of Excellence for Particle Physics at
the Terascale, Department of Physics, The University of Adelaide,
Adelaide, South Australia 5005, Australia}
\address[prague]{Institute of Particle and Nuclear Physics, Faculty of
  Mathematics and Physics, Charles University in Prague, V
  Hole\v{s}ovi\v{c}k\'{a}ch 2, 180 00 Praha 8, Czech Republic}
\address[kiasquc]{Quantum Universe Center,
Korea Institute for Advanced Study,
85 Hoegiro Dongdaemungu,
Seoul 02455, Republic of Korea}
\address[rwth]{Institute for Theoretical Particle Physics and Cosmology, RWTH Aachen University, 52074 Aachen, Germany}

  \begin{abstract}
    We document major new features and improvements of \fs
    \cite{Athron:2014yba,Athron:2017fvs}, a \mathematica and C++
    package with a dependency on the external package \sarah, that
    generates fast and precise spectrum generators.
  \end{abstract}

@@

\clearpage
\newgeometry{top=2.5cm,left=3cm,right=3cm,bottom=4cm,footskip=3em}
\section*{New version program summary}
\noindent
{\em Program title:} \fs\\[0.5em]
{\em Licensing provisions:} GPLv3\\[0.5em]
{\em Program obtainable from:} \url{http://flexiblesusy.hepforge.org/}\\[0.5em]
{\em Programming language:} C++, Wolfram/Mathematica, FORTRAN, Bourne shell\\[0.5em]
{\em Operating system:} Tested on FreeBSD, Linux, Mac OS X, Windows/Cygwin\\[0.5em]
{\em External routines:} \sarah, Boost library, Eigen library, GNU Scientific Library\\[0.5em]
{\em Optional external routines:} BLAS, Himalaya, LAPACK, LoopTools, SQLite, TSIL\\[0.5em]
{\em Journal reference of previous version:} Comput.Phys.Commun.\ 230 (2018) 145-217\\[0.5em]
{\em Does the new version supersede the previous version?:} yes \\[0.5em]
{\em Reasons for the new version:} Program extension including new models, observables and algorithms\\[0.5em]
{\em Summary of revisions:} $\mu\to e\gamma$\\[0.5em]
{\em Nature of problem:}
%
Determining the mass spectrum, mixings and further observables for an
arbitrary extension of the Standard Model, input by the user. The
generated code must find simultaneous solutions to constraints that
are specified at two or more different renormalization scales, which
are connected by renormalization group equations forming a large set
of coupled first-order differential
equations. \\[0.5em]
%
{\em Solution method:}
%
Nested iterative algorithm and numerical
minimization of the Higgs potential.\\[0.5em]
%
{\em Restrictions:}
%
The couplings must remain perturbative at all scales between the
highest and lowest boundary condition.  Tensor-like Lagrangian
parameters of rank 3 are currently not supported.  The automatic
determination of the Standard Model-like gauge and Yukawa couplings is
only supported for models that have the Standard Model gauge group
$SU(3)_C\times SU(2)_L\times U(1)_Y$ as a gauge symmetry group factor.
However, due to the modular nature of the generated code, adapting and
extending it to overcome restrictions in scope is quite straightforward.

@@

\clearpage
\section{External libraries}

\subsection{New mandatory external libraries}

Since version X.Y.Z \fs requires the following additional libraries to
be installed:
%
\begin{itemize}
\item \LoopTools: required for calculating the 1-loop decays.
\end{itemize}

\subsection{New optional external libraries}

\begin{itemize}
\item \TSIL: Since version X.Y.Z \fs can calculate 2-loop corretions
  to SM \MSbar top quark mass of the order $O(\at\as + \at^2)$ from
  Ref.~\cite{Martin:2016xsp}.  The calculation of these corrections
  requires linking the \TSIL library \cite{Martin:2005qm} to \fs.
\end{itemize}

\clearpage
\section{$\mu\to e\gamma$}
